AggregateCounts

### 2021-08-21 R/AggregateCounts.Rmd
Added the Macro Scale Region to biomass data and made plots

```{r aggregate input}
#agregate input functions
ReadAggregateFile<-function(file, species, timepoint, condition, replicate, sampleCode){#make a dataframe of the aggregate size list and add experimental parameters
  print(file)
   datos <-read.csv(file, header = F)  
   colnames(datos)<-c('Size') 
   datos<-datos %>% filter(!Size<0.5) #first filter the size
   datos<-datos %>% mutate(Species = species)#add parameters
   datos<-datos %>% mutate(TimePoint = timepoint) 
   datos<-datos %>% mutate(Condition = condition)
   datos<-datos %>% mutate(Replicate = replicate)
   datos<-datos %>% mutate(SampleCode = sampleCode) 
   datos$Species<-as.factor(datos$Species)
   datos$Replicate<-as.factor(datos$Replicate)
   datos$Condition<-as.factor(datos$Condition)
   datos$TimePoint<-as.factor(datos$TimePoint)
   datos$SampleCode<-as.factor(datos$SampleCode)
   return(datos)
}
GetAggregates<-function(){#import all experiments at once
   folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Aggregate lists')
   fileList<-list.files(path = folder, full.names = TRUE)
   growthDf<-data.frame()
   for (aggList in 1:length(fileList)){
     fileName<-str_split(basename(fileList[aggList]), "\\.")[[1]][1]#this selects the base file name and removes the ".csv"
     attrList<-str_split(fileName,"_")[[1]]
     growthDf<-rbind(growthDf,ReadAggregateFile(fileList[aggList],attrList[4],attrList[2],attrList[1],attrList[3], fileName))
     }
   return(growthDf)
}

#run functions
aggregateList<-GetAggregates()#this reads the data from Matlab
#remove bad samples and irrelevant conditions
##first remove irrelevant conditions
removable <- bind_rows(aggregateList %>% filter(Species == "Pa" & Condition == "mono"), #Pa in Sa mono culture
                       aggregateList%>% filter(Species == "Sa" & Condition == "pawt"),#Sa in all Pa mono cultures
                       aggregateList%>% filter(Species == "Sa" & Condition == "papqsL"),
                       aggregateList%>% filter(Species == "Sa" & Condition == "paphz"))
aggregateList <- anti_join(aggregateList, removable)
##then, bad samples
aggregateList<-aggregateList %>% filter(!SampleCode == "mono_d1_03_Pa" & !SampleCode == "mono_d1_03_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "phz_d4_14_Pa" & !SampleCode == "phz_d4_14_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "pqsL_d4_02_Pa" & !SampleCode == "pqsL_d4_02_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "pqsL_d4_04_Pa" & !SampleCode == "pqsL_d4_04_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "pqsL_d4_10_Pa" & !SampleCode == "pqsL_d4_10_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "wt_d4_07_Pa" & !SampleCode == "wt_d4_07_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "papqsL_d4_02_Pa" & !SampleCode == "papqsL_d4_02_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "pawt_d4_05_Pa" & !SampleCode == "pawt_d4_05_Sa")
aggregateList<-aggregateList %>% filter(!SampleCode == "phz_d4_06_Pa" & !SampleCode == "phz_d4_06_Sa")

@@

aggregateList<-aggregateList %>% mutate(Culture = recode(Condition, 'Sa mono' = "mono", 'Pa wt mono' = "mono",'pqsL mono'= "mono",'phz mono'= "mono",'wt co'= "co", 'pqsL co'= "co", 'phz co'= "co"))
aggregateList<-aggregateList%>% mutate(Strain = recode(Condition, 'Sa mono' = "Sa", 'Pa wt mono' = "wt",'pqsL mono'= "pqsL",'phz mono'= "phz",'wt co'= "wt", 'pqsL co'= "pqsL", 'phz co'= "phz"))

@@

```{r aggregate wrangling and overview plots}
#Statistics from all aggregates

#Get aggregate summary per sample
aggregateListSummaryReplicate<-aggregateList%>% group_by(Species, Strain, Culture,TimePoint, Condition, Replicate) %>% 
  summarise(Aggs = n(), biomass = sum(Size), meanAggSize = mean(Size), seAggSize = std.error(Size), maxSize = max(Size), minSize = min(Size)) %>%
  ungroup() %>% mutate(RatioMaxTotal = maxSize/biomass )

#Plot agg summary per sample
#####

ggplot()+ 
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= meanAggSize, color = Strain,shape = "MeanSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= maxSize, color = Strain, shape = "MaxSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= biomass, color = Strain, shape = "TotalBiomass"),size = 3, height = 0,width= .3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(Species~Culture, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per sample')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  scale_color_brewer(palette = "Set1")+
  theme_cowplot()+
  panel_border() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facet text
        strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))#facet background

#Get aggregate Summary per condition
aggregateListSummary<-aggregateList %>% group_by(Species, Strain, Culture,TimePoint, Condition) %>% 
  summarise(Aggs = n(), biomass = sum(Size), meanAggSize = mean(Size), seAggSize = std.error(Size), maxSize = max(Size), minSize = min(Size)) %>%
  ungroup() %>% mutate(RatioMaxTotal = maxSize/biomass ) 
#plot agg summary per condition
#####
#Version 1
ggplot()+ 
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= meanAggSize, color = Condition,shape = "MeanSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= maxSize, color = Condition, shape = "MaxSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= biomass, color = Condition, shape = "TotalBiomass"),size = 3, height = 0,width= .3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(. ~ Species, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per condition')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  scale_color_brewer(palette = "Set1")+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5),strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))
#Version 2
ggplot()+ 
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= meanAggSize, color = Strain,shape = "MeanSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= maxSize, color = Strain, shape = "MaxSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummary, aes(x = TimePoint, y= biomass, color = Strain, shape = "TotalBiomass"),size = 3, height = 0,width= .3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(Species~Culture, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per condition')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  scale_color_brewer(palette = "Set1")+
  theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5),strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))

>>>>

```{r aggregate input}
#agregate input functions
ReadAggregateFile<-function(file, species, timepoint, condition, replicate,fileName, sampleCode){#make a dataframe of the aggregate size list and add experimental parameters
  print(file)
   datos <-read.csv(file, header = F)  
   colnames(datos)<-c('Size') 
   datos<-datos %>% filter(!Size<0.5) #first filter the size
   datos<-datos %>% mutate(Species = species)#add parameters
   datos<-datos %>% mutate(TimePoint = timepoint) 
   datos<-datos %>% mutate(Condition = condition)
   datos<-datos %>% mutate(Replicate = replicate)
   datos<-datos %>% mutate(SampleCode = sampleCode)
   datos<-datos %>% mutate(FileName = fileName)
   datos$Species<-as.factor(datos$Species)
   datos$Replicate<-as.factor(datos$Replicate)
   datos$Condition<-as.factor(datos$Condition)
   datos$TimePoint<-as.factor(datos$TimePoint)
   datos$FileName<-as.factor(datos$FileName)
   datos$SampleCode<-as.factor(datos$SampleCode)

   return(datos)
}
GetAggregates<-function(){#import all experiments at once
   folder<- paste0('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/Aggregate lists')
   fileList<-list.files(path = folder, full.names = TRUE)
   growthDf<-data.frame()
   for (aggList in 1:length(fileList)){
     fileName<-str_split(basename(fileList[aggList]), "\\.")[[1]][1]#this selects the base file name and removes the ".csv"
     attrList<-str_split(fileName,"_")[[1]]
     sampleCode = paste0(attrList[1],"_",attrList[2],"_",attrList[3])
     growthDf<-rbind(growthDf,ReadAggregateFile(fileList[aggList],attrList[4],attrList[2],attrList[1],attrList[3], fileName, sampleCode))}
   return(growthDf)
}

#run functions
aggregateList<-GetAggregates()#this reads the data from Matlab

#remove bad samples and irrelevant conditions
##first remove irrelevant conditions
removable <- bind_rows(aggregateList %>% filter(Species == "Pa" & Condition == "mono"), #Pa in Sa mono culture
                       aggregateList%>% filter(Species == "Sa" & Condition == "pawt"),#Sa in all Pa mono cultures
                       aggregateList%>% filter(Species == "Sa" & Condition == "papqsL"),
                       aggregateList%>% filter(Species == "Sa" & Condition == "paphz"))
aggregateList <- anti_join(aggregateList, removable)
##then, bad samples
aggregateList<-aggregateList %>% filter(!SampleCode == "mono_d1_03")
aggregateList<-aggregateList %>% filter(!SampleCode == "phz_d4_14")
aggregateList<-aggregateList %>% filter(!SampleCode == "pqsL_d4_02")
aggregateList<-aggregateList %>% filter(!SampleCode == "pqsL_d4_04")
aggregateList<-aggregateList %>% filter(!SampleCode == "pqsL_d4_10")
aggregateList<-aggregateList %>% filter(!SampleCode == "wt_d4_07")
aggregateList<-aggregateList %>% filter(!SampleCode == "papqsL_d4_02")
aggregateList<-aggregateList %>% filter(!SampleCode == "pawt_d4_05")
aggregateList<-aggregateList %>% filter(!SampleCode == "phz_d4_06")

@@

aggregateList<-aggregateList %>% mutate(Culture = recode(Condition, 'Sa mono' = "mono", 'Pa wt mono' = "mono",'pqsL mono'= "mono",'phz mono'= "mono",'wt co'= "co", 'pqsL co'= "co", 'phz co'= "co"))
aggregateList<-aggregateList%>% mutate(Strain = recode(Condition, 'Sa mono' = "Sa", 'Pa wt mono' = "wt",'pqsL mono'= "pqsL",'phz mono'= "phz",'wt co'= "wt", 'pqsL co'= "pqsL", 'phz co'= "phz"))
#Adding the metadata
metaData <-read.csv('/Volumes/raw_data/Confocal/Carolyn/2020/Chronic wounds/images metadata.csv', header = TRUE)
metaData<-metaData %>% select(SampleCode, mouse, Position, size, Region, Zdim, xyDim,xySize)
aggregateList<-metaData %>%  merge(aggregateList)
aggregateList$Region<-as.factor(aggregateList$Region)

#---This chunk creates 'aggregateList' data frame, which contains all the aggregates per sample and it is on the correct format
```

```{r aggregate wrangling and overview plots}
#Statistics from all aggregates
#Get aggregate summary per sample
aggregateListSummaryReplicate<-aggregateList%>% group_by(SampleCode, Species, Strain, Culture,TimePoint, Condition, Replicate,Region) %>% 
  summarise(Aggs = n(), biomass = sum(Size), meanAggSize = mean(Size), seAggSize = std.error(Size), maxSize = max(Size), minSize = min(Size)) %>%
  ungroup() %>% mutate(RatioMaxTotal = maxSize/biomass)

#Plot agg summary per sample
#####
ggplot()+ 
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= meanAggSize, color = Strain,shape = "MeanSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= maxSize, color = Strain, shape = "MaxSize"),size = 3, height = 0, width = .3)+
  geom_jitter(data = aggregateListSummaryReplicate, aes(x = TimePoint, y= biomass, color = Strain, shape = "TotalBiomass"),size = 3, height = 0,width= .3)+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
  facet_grid(Species~Culture, scales = "free", space = "free") +
  ggtitle('Mean size and largest aggregate and total biomass per sample')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  scale_color_brewer(palette = "Set1")+
  theme_cowplot()+
  panel_border() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facet text
        strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))#facet background
#####
#Plot only biomass per Sample
PlotBiomasPerSample<-function(df){
ggplot()+ 
    #this is for multiple conditions
  #geom_point(data = df, aes(x = Strain, y= biomass, shape = Region, color = Culture),size = 3, position = position_dodge(width = .6))+
  geom_point(data = df, aes(x = Culture, y= biomass, color= Species, shape = Region),size = 3, 
             position = position_jitterdodge(jitter.height = 0, jitter.width = .3,dodge.width = 0.75))+
    scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)), limits = c(1, 100000000))+
      #this is for multiple conditions  
    #facet_grid(Species~TimePoint, scales = "free", space = "free") +
    #this is for only one mutant
    facet_grid(.~TimePoint, scales = "free", space = "free") +
  ggtitle('Total biomass per sample')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  scale_color_brewer(palette = "Set1")+
  scale_shape_manual(values=c(1, 2,0))+
  theme_cowplot()+
  panel_border() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facet text
        strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))#facet background
}

PlotBiomasPerSample(aggregateListSummaryReplicate)
#plotting only Sa and pa wt in mono and co-culture
PlotBiomasPerSample(aggregateListSummaryReplicate %>% filter(Condition == "Sa mono" |Condition == "wt co"|Condition == "Pa wt mono"))

#Get aggregate Summary per condition
aggregateListSummaryCondition<-aggregateListSummaryReplicate %>% group_by(Species, Strain, Culture,TimePoint, Condition, Region) %>% 
  summarise(meanAggs= mean(Aggs), meanBiomassLog = mean(log10(biomass)), SEBiomassLog = std.error((log10(biomass))), meanAggSize = mean(meanAggSize), seAggSize = std.error(meanAggSize))%>%  mutate(meanBiomass =  meanBiomassLog^10, SEBiomass = SEBiomassLog^10) 
#plot agg summary per condition
PlotBiomassPerCondition<-function(df, metric, error){
ggplot()+ 
  geom_pointrange(data = df, aes(x = Culture, y= .data[[metric]],
                                 ymax=.data[[metric]]+.data[[error]],ymin =.data[[metric]]-.data[[error]], color = Strain,shape = Region), 
             position = position_jitterdodge(jitter.height = 0, jitter.width = .3,dodge.width = 0.75))+
  scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)), limits = c(10000,1000000000))+
  facet_grid(Species~TimePoint, scales = "free", space = "free") +
  ggtitle('Mean size per condition and region')+
  labs(x = "",y = bquote(paste("Volume (", mu*"m"^3*")")))+
  #scale_shape_manual(name="Metric",values=c(MeanSize=1, MaxSize=2, TotalBiomass=17))+
  scale_color_brewer(palette = "Set1")+
    theme_cowplot()+
  theme(plot.title = element_text(hjust = 0.5),strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facets
          strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))
}
PlotBiomassPerCondition(aggregateListSummaryCondition,"meanBiomass", "SEBiomass")
PlotBiomassPerCondition(aggregateListSummary,"biomass")
#combining all 
PlotBiomasPerSampleAndConditionWT<-function(df1,df2){
ggplot()+ 
    geom_point(data = df1, aes(x = Culture, y= meanBiomassLog,color = Species,shape = Region), 
             position = position_jitterdodge(jitter.height = 0, jitter.width = .3,dodge.width = 0.75), size = 4)+
  geom_point(data = df2, aes(x = Culture, y= log10(biomass), color= Species, shape = Region),size = 1, 
             position = position_jitterdodge(jitter.height = 0, jitter.width = .3,dodge.width = 0.75))+
    #scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)), limits = c(1, 100000000))+
      facet_grid(.~TimePoint, scales = "free", space = "free") +
  ggtitle('Biomass per sample')+
  labs(x = "",y = bquote(paste("Volume (log10", mu*"m"^3*")")))+
  scale_color_brewer(palette = "Set1")+
  scale_shape_manual(values=c(16, 1))+
  ylim(0,8)+
    theme_cowplot()+
  panel_border() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facet text
        strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))#facet background
}
PlotBiomasPerSampleAndConditionWT(aggregateListSummaryCondition%>% filter(Condition == "Sa mono" |Condition == "wt co"|Condition == "Pa wt mono"),
                                aggregateListSummaryReplicate %>% filter(Condition == "Sa mono" |Condition == "wt co"|Condition == "Pa wt mono"))
PlotBiomasPerSampleAndCondition<-function(df1,df2){
ggplot()+ 
    geom_point(data = df1, aes(x = Culture, y= meanBiomassLog,color = Strain,shape = Region), 
             position = position_jitterdodge(jitter.height = 0, jitter.width = .3,dodge.width = 0.75), size = 4)+
  geom_point(data = df2, aes(x = Culture, y= log10(biomass), color= Strain, shape = Region),size = 1, 
             position = position_jitterdodge(jitter.height = 0, jitter.width = .3,dodge.width = 0.75))+
    #scale_y_log10( breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)), limits = c(1, 100000000))+
      facet_grid(Species~TimePoint, scales = "free", space = "free") +
  ggtitle('Biomass per sample')+
  labs(x = "",y = bquote(paste("Volume (log10", mu*"m"^3*")")))+
  scale_color_brewer(palette = "Set1")+
  scale_shape_manual(values=c(4,16, 1))+
  ylim(0,8)+
    theme_cowplot()+
  panel_border() +
  theme(plot.title = element_text(hjust = 0.5), 
        strip.text = element_text(size = 11, face = "italic",margin = margin(.1,0,.1,0, "cm")),#this is for the facet text
        strip.background = element_rect(color="black", fill="white", size=1, linetype="solid"))#facet background
}
PlotBiomasPerSampleAndCondition(aggregateListSummaryCondition,aggregateListSummaryReplicate)

@@
